Geometric variational problems

### 2020-11-14 NSF2020/Description/newspecificaim3.tex
yny 1114

We include a background flow by replacing the third equation in
\eqref{eq:stokes} with the condition $\mathbf{u}(\mathbf{x})
\to \mathbf{u}_{\infty}(\mathbf{x})$ as $|\mathbf{x}| \to
\infty$. To incorporate the far-field flow, we use the representation 
\begin{align}
\label{PowerMiranda}
  {\bf u} = {\bf u}_{\infty} + K\boldsymbol{\eta} + 
    \sum_{i=1}^N \left(S(\cdot, {\bf a}_i) {\bf F}_i + 
                 R(\cdot, {\bf a}_i) {\bf G}_i\right),
\end{align}
where $S$ and $R$ are stokeslets and rotlets supported at the respective
particle centers~\cite{leal_2007} and $K\boldsymbol{\eta}$ is a layer
potential for the unknown density function $\boldsymbol{\eta}$. With the
exception of the rigid body conditions, the
representation~\eqref{PowerMiranda} automatically satisfies all the
hydrodynamic equations including the far-field condition. The rigid
motion conditions follow by requiring the viscous stress vanishes across
the particle boundaries.

We will compare the behavior of our particle-based vesicles in Stokes
flow to well-established models that enforce area incompressibility and
volume conservation in vesicle membrane
dynamics~\cite{torres-sanchez_millan_arroyo_2019,
mahapatra_saintillan_rangamani_2020, Steigmann99, C6SM02452A}. Our
preliminary results show that the particle-based bilayers have the same
large area modulus as real lipid bilayer membranes, so we expect
realistic results from our flow simulations. Under moderate shear rates,
the vesicle elongates and takes on the tank-treading elliptical shape.
We can directly check for area compressibility and volume conservation
properties from particle simulations by tracking the change in area
density of the midplane as a function of shear flow rate, along with
evaluating the tangential divergence of the velocity.

Bilayer membranes have a small permeability to
water~\cite{323e9a2f0c58487ea82518d7a1f96485}, and modelers often assume
a vesicle membrane that conserves volume. In our particle-based
approach, water motion across the bilayer is limited by the size of the
gaps between the particles, which is an artifact of coarse-graining.
Making these gaps smaller comes at the expense of numerical accuracy,
and we will assess if there is a reasonable trade-off between
approximate volume conservation and simulation complexity.

Finally, the particle-based vesicle bilayers have two distinct leaflets
consisting of the particles in contact with the vesicle interior, and
those in contact with the surrounding fluid. Since there is a nonzero
separation between these layers, an important effect we observe is that
the leaflets slide against one another under shear flow. This implies
that part of the viscous dissipation in the aqueous phase is enhanced by
intermonolayer friction~\cite{SHKULIPA2005823, ShkulipaThesis}.
Intermonolayer slip enters zero-thickness membrane models as a velocity
jump boundary condition and leads to multiple solution
states~\cite{schwalbe_vlahovska_miksis_2010}. In the present proposal,
intermonolayer slip is an artifact of monolayer independence and we can
control for slip as a function of shear rate, vesicle diameter, and
particle geometry.

>>>>

We include a background flow by replacing the third equation in
\eqref{eq:stokes} with the condition $\mathbf{u}(\mathbf{x})
\to \mathbf{u}_{\infty}(\mathbf{x})$ as $|\mathbf{x}| \to
\infty$. To incorporate the far-field flow, we use the representation 
\begin{align}
\label{PowerMiranda}
  {\bf u} = {\bf u}_{\infty} + K\boldsymbol{\eta} + 
    \sum_{i=1}^N \left(S(\cdot, {\bf a}_i) {\bf F}_i + 
                 R(\cdot, {\bf a}_i) {\bf G}_i\right),
\end{align}
where $S$ and $R$ are stokeslets and rotlets supported at the respective
particle centers~\cite{leal_2007} and $K\boldsymbol{\eta}$ is a layer
potential for the unknown density function $\boldsymbol{\eta}$. With the
exception of the rigid body conditions, the
representation~\eqref{PowerMiranda} automatically satisfies all the
hydrodynamic equations including the far-field condition. The rigid
motion conditions follow by requiring the viscous stress vanishes across
the particle boundaries.
%
We will compare the behavior of our particle-based vesicles in Stokes
flow to well-established models that enforce area incompressibility and
volume conservation in vesicle membrane
dynamics~\cite{torres-sanchez_millan_arroyo_2019,mahapatra_saintillan_rangamani_2020, Steigmann99, C6SM02452A}.
Some preliminary results are in \S \ref{sec:preliminary_work}. 
%
%Our
%preliminary results show that the particle-based bilayers have the same
%large area modulus as real lipid bilayer membranes, so we expect
%realistic results from our flow simulations. Under moderate shear rates,
%the vesicle elongates and takes on the tank-treading elliptical shape.
%We can directly check for area compressibility and volume conservation
%properties from particle simulations by tracking the change in area
%density of the midplane as a function of shear flow rate, along with
%evaluating the tangential divergence of the velocity.
%
%Bilayer membranes have a small permeability to
%water~\cite{323e9a2f0c58487ea82518d7a1f96485}, and modelers often assume
%a vesicle membrane that conserves volume. In our particle-based
%approach, water motion across the bilayer is limited by the size of the
%gaps between the particles, which is an artifact of coarse-graining.
%Making these gaps smaller comes at the expense of numerical accuracy,
%and we will assess if there is a reasonable trade-off between
%approximate volume conservation and simulation complexity. 
%
%Finally, the particle-based vesicle bilayers have two distinct leaflets
%consisting of the particles in contact with the vesicle interior, and
%those in contact with the surrounding fluid. Since there is a nonzero
%separation between these layers, an important effect we observe is that
%the leaflets slide against one another under shear flow. This implies
%that part of the viscous dissipation in the aqueous phase is enhanced by
%intermonolayer friction~\cite{SHKULIPA2005823, ShkulipaThesis}.
%Intermonolayer slip enters zero-thickness membrane models as a velocity
%jump boundary condition and leads to multiple solution
%states~\cite{schwalbe_vlahovska_miksis_2010}. In the present proposal,
%intermonolayer slip is an artifact of monolayer independence and we can
%control for slip as a function of shear rate, vesicle diameter, and
%particle geometry.

### 2020-11-14 NSF2020/Description/preliminarywork.tex
yny 1114

\begin{wraptable}[8]{r}{0.55\textwidth}
\centering{
  \begin{tabular}{|l|r|r|p{0.1\textwidth}|}
    \hline
          & HAP                     & experiment          & ref.\\
    \hline 
    $\KB$ & $11 \pm 2.5$ \kBT       & 10 \kBT               & \cite{Naetal15,VeBrPa15,NAGLE2000159,PhysRevLett.113.248102}\\
    \hline 
    $\KA$ & $34$ \kBT \; nm$^{-2}$  & 30--40 \kBT nm$^{-2}$ & \cite{Nagle17, Nagle17-2}\\
    \hline 
    $\KTH$ & $12$ \kBT \; nm$^{-2}$ & 10 \kBT nm$^{-2}$     & \cite{KUZMIN2005, KoNa15} \\ \hline
  \end{tabular}
}
  \vspace{-8pt}
\caption{\label{tab:moduli} \footnotesize Comparision of values elastic
  moduli from the experimental literature and values derived by HAP
  simulation.} 
\end{wraptable}
\section{Preliminary work}
\label{sec:preliminary_work}
The HAP project was initiated in 2017 by PI RR and YNY. The lead author
on our first paper~\cite{Fu2018_SIAM} was Szu-Pei Fu, who was PI YNY's
PhD student and is currently a postdoc with PI RR. The collaboration
involved theoretical work and simulation. The theoretical work included
model development, proving the first variation formula~\eqref{stress}
and force-free law, and an energy-based uniqueness theorem for the
exterior domain problem~\eqref{SL}. Coauthors Andreas Kl\"ockner
(Computer Science, UIUC, NSF CAREER \#1654756) and his PhD student Matt
Wala provided their Quadrature by Expansion, or QBX algorithm and Fu ran
simulations on NJIT clusters.

\begin{wraptable}[11]{l}{0.42\textwidth}
\centering{
  \begin{tabular}{|l|l|l|l|}
    $n$   & (rel.~err.)$_F$
    &  (rel.~err.)$_G$ \\
    \hline
% 32   &   4.07466     &   1.56437\\
% 64   &   0.24664     &   0.03476\\
% 128  &   0.00063     &   0.00006\\
  32  & $4.07 \times 10^{+0}$ & $1.56 \times 10^{+0}$ \\
  64  & $2.47 \times 10^{-1}$ & $3.48 \times 10^{-2}$ \\
  128 & $6.30 \times 10^{-4}$ & $6.00 \times 10^{-5}$ 
  \end{tabular}}
\vspace{-5pt}
\caption{\label{tab:spectral_force} 
\footnotesize  Relative numerical errors (rel.~err.)$_F = \max_{i}
  \|\mathbf{F}_i-\mathbf{F}_i^{\text{exact}}\|/\|\mathbf{F}_i^{\text{exact}}\|$
  and (rel.~err.)$_G = \max_{i}
  \|\mathbf{G}_i-\mathbf{G}_i^{\text{exact}}\|/\|\mathbf{G}_i^{\text{exact}}\|$
  for force and torque respectively as a function of number of grid
  points $n$ per particle.  The data is for $N = 5$ particles; two of
  the particles are nearly touching at a distance 1\% of particle
  diameter.} 
\end{wraptable}
The simulation results were encouraging. Small particle-number
simulations successfully demonstrated gradient-driven self-assembly of
amphiphilic particles into two-dimensional micelles and bilayer membranes.
The model parameters were the decay length $\rho$, interfacial tension
$\gamma$, particle geometry, and repulsion strength. The particle
diameters were the physical monolayer thickness, 2.0 nm~\cite{Boal}, and
we chose a repulsion strength that gave adequate interparticle spacing.
For the decay length, we used $\rho = 2.5$ nm based on measurements of
hydrophobic attraction between surfactant-coated
membranes~\cite{Eriksson1989, Lin2005, Parsegian, Israelachvili80}. For
the interfacial tension, we used $\gamma = 4$ pN nm$^{-1}$ based on
measurements for single-component bilayer lipid
membranes~\cite{GarciaSaez, KUZMIN2005, Petelska2012, Jackson2016}.

\begin{wrapfigure}[]{r}{0.5\textwidth}
  \centering{\includegraphics[width=0.5\textwidth]{Figures/TankTreading.pdf}}
  \vspace{-25pt}
  \caption{\label{fig:tank_treading} \footnotesize
    Panel A shows water activity and flow direction for  a two-dimensional
    particle-based vesicle in a shear flow; $\mu  = 1$ mPa s,
    $\dot{\gamma} = 0.5$ $\mu$s$^{-1}$. Panel B plots the
    mean angular velocities for the outer leaflet and the inner
    respectively.  Panel C shows that the trajectory of the particle
    centers forms two ellipses with well-defined inclination angle
    $\Theta$. Panel D shows the evolution of the inclination angle.
    Panels B and D share the same time axis.}
\end{wrapfigure}
The paper \cite{Fu2018_SIAM} considered three types of simulations. The first measured bending by
loading a partially clamped planar bilayer. The second used a harmonic
bond to dilate a circular bilayer. The third measured tilt using a decay
equation from~\cite{KUZMIN2005}. These simulations isolated three of the
five deformations of~\eqref{ansatz3}, enabling us to read off elastic
moduli from simulation data. The results agreed remarkably well with the
values reported in the experimental literature (Table~\ref{tab:moduli}).
The agreement is underscored by the fact that the two main HAP
parameters, $\gamma$ and $\rho$, were assigned physical values from the
outset rather than being tuned to fit data.

The PI BQ, RR and YNY initiated a collaboration in Summer 2020 to
develop the hydrodynamic model of \S\ref{subsec:specific_aim_3}.  To
date, we have written solvers for the two-dimensional boundary integral
formulations of~\eqref{SL} and~\eqref{eq:stokes} that achieve
third-order accuracy up to the boundary. The specialized reciprocal
identity~\eqref{eq:reciprocal} gives spectral accuracy for force and
torque (Table~\ref{tab:spectral_force}).

@@

\begin{wrapfigure}[11]{r}{0.22\textwidth}
  \centering{\includegraphics[width=0.22\textwidth]{Figures/PW_fig5.pdf}}
  \caption{\label{fig:flow_rupture} \footnotesize Rupture of a
  two-dimensional vesicle at large shear rates.}
\end{wrapfigure}
The above considerations further justify using the particle-based model
to study material properties as outlined in
\S\ref{subsec:specific_aim_1}. Furthermore, the particle system allows
for intermonolayer slip~\cite{SHKULIPA2005823, ShkulipaThesis}. In
Figure \ref{fig:tank_treading}B, the inner angular velocity is slightly
less than the outer angular velocity. Intermonolayer slip enters
zero-thickness membrane models as a velocity jump boundary condition
\cite{schwalbe_vlahovska_miksis_2010}, for example, but in the present
model, it is a consequence of monolayer independence. For large shear
rates, viscous forces exceed the hydrophobic attraction and the vesicle
ruptures; the flow carries away segmented membrane patches (Figure
\ref{fig:flow_rupture}).

>>>>

\begin{wraptable}[8]{r}{0.55\textwidth}
\centering{
  \begin{tabular}{|l|r|r|p{0.1\textwidth}|}
    \hline
          & HAP                     & experiment          & ref.\\
    \hline 
    $\KB$ & $11 \pm 2.5$ \kBT       & 10 \kBT               & \cite{Naetal15,VeBrPa15,NAGLE2000159,PhysRevLett.113.248102}\\
    \hline 
    $\KA$ & $34$ \kBT \; nm$^{-2}$  & 30--40 \kBT nm$^{-2}$ & \cite{Nagle17, Nagle17-2}\\
    \hline 
    $\KTH$ & $12$ \kBT \; nm$^{-2}$ & 10 \kBT nm$^{-2}$     & \cite{KUZMIN2005, KoNa15} \\ \hline
  \end{tabular}
}
  \vspace{-8pt}
\caption{\label{tab:moduli} \footnotesize Comparision of values elastic
  moduli from the experimental literature and values derived by HAP
  simulation.} 
\end{wraptable}
\section{Preliminary work}
\label{sec:preliminary_work}
PIs RR and YNY started work on HAP modeling with Szu-Pei Fu (then PI YNY's student and now PI RR's postdoc) and collaborators in 2017. 
%
%The HAP project was initiated in 2017 by PI RR and YNY. The lead author
%on our first paper~\cite{Fu2018_SIAM} was Szu-Pei Fu, who was PI YNY's
%PhD student and is currently a postdoc with PI RR. The collaboration
%involved theoretical work and simulation. The theoretical work included
%model development, proving the first variation formula~\eqref{stress}
%and force-free law, and an energy-based uniqueness theorem for the
%exterior domain problem~\eqref{SL}. 
%Coauthors Andreas Kl\"ockner
%(Computer Science, UIUC, NSF CAREER \#1654756) and his PhD student Matt
%Wala provided their Quadrature by Expansion, or QBX algorithm and Fu ran
%simulations on NJIT clusters.
%
\begin{wraptable}[11]{l}{0.42\textwidth}
\centering{
  \begin{tabular}{|l|l|l|l|}
    $n$   & (rel.~err.)$_F$
    &  (rel.~err.)$_G$ \\
    \hline
% 32   &   4.07466     &   1.56437\\
% 64   &   0.24664     &   0.03476\\
% 128  &   0.00063     &   0.00006\\
  32  & $4.07 \times 10^{+0}$ & $1.56 \times 10^{+0}$ \\
  64  & $2.47 \times 10^{-1}$ & $3.48 \times 10^{-2}$ \\
  128 & $6.30 \times 10^{-4}$ & $6.00 \times 10^{-5}$ 
  \end{tabular}}
\vspace{-5pt}
\caption{\label{tab:spectral_force} 
\footnotesize  Relative numerical errors (rel.~err.)$_F = \max_{i}
  \|\mathbf{F}_i-\mathbf{F}_i^{\text{exact}}\|/\|\mathbf{F}_i^{\text{exact}}\|$
  and (rel.~err.)$_G = \max_{i}
  \|\mathbf{G}_i-\mathbf{G}_i^{\text{exact}}\|/\|\mathbf{G}_i^{\text{exact}}\|$
  for force and torque respectively as a function of number of grid
  points $n$ per particle.  The data is for $N = 5$ particles; two of
  the particles are nearly touching at a distance 1\% of particle
  diameter.} 
\end{wraptable}
The results are summarized in \cite{Fu2018_SIAM}, where
small particle-number simulations successfully demonstrated gradient-driven self-assembly of
amphiphilic particles into two-dimensional micelles and bilayer membranes with realistic values of
the model parameters:  decay length $\rho=2.5$ nm (based on measurements of
hydrophobic attraction between surfactant-coated
membranes~\cite{Eriksson1989, Lin2005, Parsegian, Israelachvili80}),
particle diameter 2.0 nm (the physical monolayer thickness, 2.0 nm~\cite{Boal}),
and interfacial tension $\gamma = 4$ pN nm$^{-1}$ based on
measurements for single-component bilayer lipid
membranes~\cite{GarciaSaez, KUZMIN2005, Petelska2012, Jackson2016}.

%\begin{wrapfigure}[]{r}{0.5\textwidth}
%  \centering{\includegraphics[width=0.5\textwidth]{Figures/TankTreading.pdf}}
%  \vspace{-25pt}
%  \caption{\label{fig:tank_treading} \footnotesize
%    Panel A shows water activity and flow direction for  a two-dimensional
%    particle-based vesicle in a shear flow; $\mu  = 1$ mPa s,
%    $\dot{\gamma} = 0.5$ $\mu$s$^{-1}$. Panel B plots the
%    mean angular velocities for the outer leaflet and the inner
%    respectively.  Panel C shows that the trajectory of the particle
%    centers forms two ellipses with well-defined inclination angle
%    $\Theta$. Panel D shows the evolution of the inclination angle.
%    Panels B and D share the same time axis.}
%\end{wrapfigure}
The paper \cite{Fu2018_SIAM} considered three types of simulations. The first measured bending by
loading a partially clamped planar bilayer. The second used a harmonic
bond to dilate a circular bilayer. The third measured tilt using a decay
equation from~\cite{KUZMIN2005}. These simulations isolated three of the
five deformations of~\eqref{ansatz3}, enabling us to read off elastic
moduli from simulation data. The results agreed remarkably well with the
values reported in the experimental literature (Table~\ref{tab:moduli}).
The agreement is underscored by the fact that the two main HAP
parameters, $\gamma$ and $\rho$, were assigned physical values from the
outset rather than being tuned to fit data.

PIs BQ, RR and YNY initiated a collaboration in Summer 2020 to
develop the hydrodynamic model of \S\ref{subsec:specific_aim_3}.  To
date, we have written solvers for the two-dimensional boundary integral
formulations of~\eqref{SL} and~\eqref{eq:stokes} that achieve
third-order accuracy up to the boundary. The specialized reciprocal
identity~\eqref{eq:reciprocal} gives spectral accuracy for force and
torque (Table~\ref{tab:spectral_force}).
\begin{wrapfigure}[]{r}{0.5\textwidth}
  \centering{\includegraphics[width=0.5\textwidth]{Figures/TankTreading.pdf}}
  \vspace{-25pt}
  \caption{\label{fig:tank_treading} \footnotesize
    Panel A: A two-dimensional
    particle-based vesicle in a shear flow (arrows); $\mu  = 1$ mPa s,
    $\dot{\gamma} = 0.5$ $\mu$s$^{-1}$. The color coding is for HAP. Panel B: The
    mean angular velocities for the outer leaflet and the inner
    respectively.  Panel C: The trajectory of the particle
    centers forms two ellipses with well-defined inclination angle
    $\Theta$. Panel D: The evolution of the inclination angle.
    Panels B and D share the same time axis.}
\end{wrapfigure}

@@

\begin{wrapfigure}[11]{l}{0.22\textwidth}
  \centering{\includegraphics[width=0.22\textwidth]{Figures/PW_fig5.pdf}}
  \caption{\label{fig:flow_rupture} \footnotesize Rupture of a
  two-dimensional vesicle at large shear rates.}
\end{wrapfigure}
The above considerations further justify using the particle-based model
to study material properties as outlined in
\S\ref{subsec:specific_aim_1}. Furthermore, the particle system allows
for intermonolayer slip~\cite{SHKULIPA2005823, ShkulipaThesis}. In
Figure \ref{fig:tank_treading}B, the inner angular velocity is slightly
less than the outer angular velocity. Intermonolayer slip enters
zero-thickness membrane models as a velocity jump boundary condition
\cite{schwalbe_vlahovska_miksis_2010}, for example, but in the present
model, it is a consequence of monolayer independence. For large shear
rates, viscous forces exceed the hydrophobic attraction and the vesicle
ruptures; the flow carries away segmented membrane patches (Figure
\ref{fig:flow_rupture}).
